### 2024-11-15 paper/nature-scientific-data-response-letter/response.tex

\begin{figure}[h]
    \centering
    \includegraphics[width=0.65\textwidth]{figures/fig_example_clustering.png}
    \caption{Example for clustering of buses. Taken from the interactive map linked on Zenodo. \cite{xiongPrebuiltElectricityNetwork2024}}
    \label{fig:example_clustering}
\end{figure}

Excerpt of the clustering algorithm as included in the revised manuscript: From each substation polygon and each virtual bus (endpoint of lines/cables), a buffer of 500 meters is created. Substations and virtual buses are then clustered if their buffers overlap (darkred polygon). Then an internal point based on the `Pole of Inaccessibility' \cite{garcia-castellanosPolesInaccessibilityCalculation2007} is determined. With this methodology, we can find a polygon-internal point, even if the polygon is non-convex. If the clustered area contained one or multiple substation(s) from the OSM data, the largest substation sets the geographical coordinates (bright red point) of the clustered substation (yellow polygon). This ensures, that the obtained coordinates are always within the perimeters of an original substation. Polygon internal elements are removed and line/cable endings are modified to exactly end in the determined internal point. Transformers and converters are finally added if necessary and accordingly (see manuscript).

\RC What happens to this model if overpass turbo is not maintained anymore?

\AR Thank you for your question. We have considered this scenario --- in the event that Overpass turbo is no longer maintained, we can either adapt to another API or use so-called \textit{planet} files, that OpenStreetMap directly provides (\href{https://planet.openstreetmap.org}{https://planet.openstreetmap.org}). These files, provided in .pbf format, are updated every week and contain snapshots of the entire OSM database. The PyPSA-Earth model is using these files via their developed \textit{earth-osm} Python library (\href{https://github.com/pypsa-meets-earth/earth-osm}{https://github.com/pypsa-meets-earth/earth-osm}). Per default, we have deliberately chosen not to use \textit{planet} files in our implementation, as they are very large and require a lot of computational resources and time to process. Electricity-related layers only make up a small fraction of the entire OSM database, this is why we use Overpass Turbo to extract only the relevant data. Further, we would like to highlight that Overpass turbo is a well-established and widely used tool, and we are confident that suitable alternatives would be available in such a scenario. Anyway, we notified maintainers of \textit{earth-osm} repository to consider supporting the Overpass API option, which could enhance backup options and improve the tool's overall resilience; for details see \href{https://github.com/pypsa-meets-earth/earth-osm/issues/54}{https://github.com/pypsa-meets-earth/earth-osm/issues/54}. If needed, users can even decide to create their own OSM database mirror and Overpass turbo client if they wish, as the OSM wiki describes (\href{https://wiki.openstreetmap.org/wiki/Overpass_API/Installation}{https://wiki.openstreetmap.org/wiki/Overpass\_API/Installation}).
